---
title: "359HW5"
author: "Mikayla Mulgrew V00923052"
date: "05/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Q1.
```{r Q1}
beef<-read.table(file='C:/Users/Mik/Documents/UVIC/STAT359/data/data/beef.txt', sep=""
                 , header=TRUE, na.strings="NA")
attach(beef)

library(knitr)
kable(beef, caption="Beef Consumption Data", align='r')

#20 data points, therefore want no more than 6 parameters

pairs(beef, panel=panel.smooth)
#seems to be a positive correlation price and income and price and year, income and year
#a few graphs have large dips in points occurring around the 30's
#seeing some effects of the great depression?
#doesn't seem to be any correlation between price and beef consumption
#there may be some correlation between income and beef consumption
#may be some negative/curved correlation between pork consumption and price

beef.model<-lm(Beef.Consumption~Income*Year*Price*Pork.Consumption)
summary(beef.model)
b1.model<-update(beef.model,~.-Income:Year:Price:Pork.Consumption)
summary(b1.model)
b2.model<-update(b1.model,~.-Income:Year:Price)
summary(b2.model)
b3.model<-update(b2.model,~.-Income:Year:Pork.Consumption)
summary(b3.model)
b4.model<-update(b3.model,~.-Income:Price:Pork.Consumption)
summary(b4.model)
b5.model<-update(b4.model,~.-Year:Price:Pork.Consumption)
summary(b5.model)
b6.model<-update(b5.model,~.-Price:Pork.Consumption)
summary(b6.model)
b7.model<-update(b6.model,~.-Year:Pork.Consumption)
summary(b7.model)
b8.model<-update(b7.model,~.-Income:Pork.Consumption)
summary(b8.model)
b9.model<-update(b8.model,~.-Income:Price)
summary(b9.model)
b10.model<-update(b9.model,~.-Income:Year)
summary(b10.model)

#Went from 6 parameters to 5 because one of the 6 terms still was not significant.

par(mfrow=c(2,2))
plot(b8.model,which=c(1,2,4))
```

The variance of the residuals is very consistent.
The QQ-plot gives some indication of normality. 
The interaction between Year and Price is somewhat significant.
Income and Pork-COnsumption are the most significant in the model, so they contribute
the most to the variability of beef consumption. 
This model explains approximately 94.26% of the variability in beef consumption.

Q2.
```{r Q2}
cholesterol<-read.table(file='C:/Users/Mik/Documents/UVIC/STAT359/data/data/chol.txt', 
                        sep="", header=TRUE)
attach(cholesterol)

library(knitr)
kable(cholesterol, caption="Cholesterol Study", align='r')

plot(BMI, CHOL, xlab='Body Mass Index (BMI)', ylab='Cholesterol Level')
title('BMI and Cholesterol Level')
#appears to be a positive correlation. Higher Cholesterol level for higher BMI, with some exceptions. 
#some observations have much higher cholesterol levels than other BMI's at the same level. 
model1<-lm(CHOL~BMI)
abline(model1)
summary(model1)
#with an R-squared value of only 0.2865, model1 is not a good fit
#with a p-value of 0.002303, reject H0 = linear model is a good fit/no relationship between 
#Cholesterol and BMI.

2*(1-pt(3.353,length(BMI)-2))
#value = 0.002305173

coef(model1)[2]
#value =0.3089659 

confint(model1)

pairs(cholesterol, panel=panel.smooth)
##appears to be some positive correlation between cholesterol and age. and maybe 
#between BMI and cholesterol
#Age and BMI don't seem to have much correlation, very scattered

model2<-lm(CHOL~BMI*Age)
summary(model2)
model3<-update(model2,~.-BMI:Age)
summary(model3)


model4<-step(model3)
summary(model4)
par(mfrow=c(2,2))
plot(model4)

#The residual plots show that the variance and normality are reasonable 
#The QQ-plot seems to have a heavier right side, will try log transform

model5<-lm(log(CHOL)~BMI+Age)
model6<-step(model5)
summary(model6)
par(mfrow=c(2,2))
plot(model6)
```

First tested the model CHOL~BMI without Age. But with an R-squared value of 0.2865, pursued a more comprehensive model. But the p-value is 0.002303 so there is evidence of a relationship between Cholesterol and BMI. Adding Age to the model, Next testing BMI*Age to get the interaction term as well. The interaction between BMI was not significant so the model was updated and ends up being BMI+Age. After plotting the residual plots for the minimally adequate model, there was a reasonable amount of variance but there is a heavier right tail on the QQ-plot. Performing a log transform seemed more normal with the exception of point 24. The R-squared value for the log transformed data was also marginally better than previous model suggesting a better explanation of variability. Points 24,25,17 are extreme in each of the plots. Since the effects of BMI and Age are additive, there interaction does not have a significant relationship with Cholesterol level.
